\Large{All in the family:\\German twin finishing times in\\the 2016 women's Olympic marathon}

### 2017-02-17 paper/chance_memo.tex
ref memo start

\noindent
\begin{tabular}{ll}
  \textbf{To:} &  Scott Evans, editor, \emph{Chance}\\
  \textbf{From:} & David Cottrell and Michael C.\ Herron, Dartmouth College\\
  \textbf{Date:} & \today \\
  \textbf{Re:} &  Revisions to women's marathon submission
\end{tabular}

@@

This memorandum is a response to your February 16, 2017, email
regarding our manuscript, ``All in the family: German twin finishing
times in the 2016 women's Olympic marathon.'' Thank you very much for
inviting us to submit a revision of our manuscript, and this
memorandum explains how we have updated it in light of comments from
you and the Chance Editorial Board.

blah blah blah

>>>>

\noindent
\begin{tabular}{ll}
  \textbf{To:} &  Scott Evans, editor, \emph{CHANCE}\\
  \textbf{From:} & David Cottrell and Michael C.\ Herron, Dartmouth College\\
  \textbf{Date:} & \today \\
  \textbf{Re:} &  Revisions to women's marathon submission
\end{tabular}

@@

This memorandum is a response to your February 16, 2017, email
regarding our manuscript, ``All in the family: German twin finishing
times in the 2016 women's Olympic marathon.'' Thank you very much for
inviting us to submit a revision of our manuscript to \emph{CHANCE},
and this memorandum explains how we have updated it in light of
comments from you and the Editorial Board.

\begin{itemize}
\item ``Am not convinced of quadratic model. Two points in upper right
  may have influenced this. What is p-value for linear vs quadratic?
  What happens if those 2 points are removed? Could also simply plot
  differences from personal best as frequency dist'n. Could also plot
  projected from half vs actual.''

\item ``I printed this out in black and white. Would help see them if
  Hahners were darkest color.''

\item ``a similar conclusion is reached?''

  We changed the language in the paper so it uses this phrasing.

\item ``pairs are not completely independent as same runner can be in
  different pairs. Not as though runners are drawn with replacement.''

\item ``due to subsequent DNF's?''

  We changed the language in the paper so it uses this phrasing.

\item ``I don't get this part. If Yi is function of Xi, personal best,
  won't estimates be same for each simulation?  What I would try for
  an analysis would be look at distribution of finishing times,
  expecting it to be normal, as most human performance is fairly
  normal (most are average, some above, some below). I'd use the
  observed mean and variance of the finishers plus a random error, and
  see what this theoretical dist'n looks like. How often are finishers
  within x seconds of each other? Could also use historical marathon
  data from other races.''

\item ``5a'' and ``5b''

\item ``Would help to expand left side bins, as these are the finishes
  of main interest.''

\item ``what are interval widths in seconds? Hard to see.''

\item ``Suddenly Latin?''

  We have changed the Latin to, all things equal.

\item ``likely''

\end{itemize}

### 2017-02-17 paper/paper.tex
ref memo start

The cumulative residuals depicted in the two panels of Figure
\ref{fig:residuals} are consistent with our interpretation of the
Hahner twin's Olympic finishing times as not particularly remarkable.
With personal best time as a measure of marathon talent as in Figure
\ref{fig:studentizedresiduals}, the two Hahner residuals are in the
right tail of the residual distribution; however, their locations are
not extreme.  One residual is located at the 88th percentile and the
other at the 93rd.  While these two residuals are in the tail end of
the residual distribution, they are not the major outliers that would
lead us to think that the Hahner twin's marathon finishing times were
extremely slow.  Moreover, the residuals for the Kim twins are
similarly not remarkable, although they are in the left tail of the
residual distribution, indicating that the Kims ran faster than one
would have expected.  If one uses half marathon split times as a
measure of marathon talent, as in Figure
\ref{fig:studentizedresiduals_half}, similar results obtain.  Neither
Hahner twin had a finishing time that was particularly unusual given
her half marathon split, and this applies to the Kim twins as well.

@@

Figure \ref{fig:secondsbehind} provides a sense of the lockstep nature
of the Kim twins, and it describes each Olympic runner's status at
various split times on the marathon course.  Each dot in the figure
depicts a recorded split and the number of seconds each runner was
behind the race leader at the time.  There are more dots at earlier
splits due to the accumulation of DNFs.  The usual color scheme
applies here, and the Estonian triplet DNF prior to the half marathon
split is evident. As noted in the figure, Lisa Hahner was ahead of her
sister through 15 kilometers at which point Anna took the lead.
Figure \ref{fig:secondsbehind} contains two red dots representing the
North Korean Kim twins, but this is not visually apparent because the
Kim twins had identical split times during the entire marathon.

@@

An important point regarding our simulations is the way that we
handled DNFs. As our description above indicates, we assumed that DNF
probabilities are the same for all runners and that the likelihood of
a DNF is a not function of a runner's anticipated marathon finishing
time. The rug marks in Figure \ref{fig:45degreeplot} suggest that
runners with better personal best times may be more likely to DNF than
other runners, \emph{ceteris paribus}. We suspect that this occurs
because some better runners may expend excessive energy trying to
achieve a good result in the marathon and in so doing injure or
exhaust themselves; lesser runners, in contrast, may be content to
finish respectably. Regardless of the validity of this conjecture,
Figure \ref{fig:45degreeplot} shows that the Hahner twins are
representative of the sort of runners who DNFed in Rio. Since our
simulated Hahner statistics are conditioned on both Hahner twins
finishing, it follows that they are conservative. The fact that both
women finished the marathon was notable in and of itself, and by
discounting the possibility of a Hahner DNF we are giving the benefit
of the doubt to the Hahner twins.

>>>>

The cumulative residuals depicted in the two panels of Figure
\ref{fig:residuals} are consistent with our interpretation of the
Hahner twin's Olympic finishing times as not particularly remarkable.
With personal best time as a measure of marathon talent as in Figure
\ref{fig:studentizedresiduals}, the two Hahner residuals are in the
right tail of the residual distribution; however, their locations are
not extreme.  One residual is located at the 88th percentile and the
other at the 93rd.  While these two residuals are in the tail end of
the residual distribution, they are not the major outliers that would
lead us to think that the Hahner twin's marathon finishing times were
extremely slow.  Moreover, the residuals for the Kim twins are
similarly not remarkable, although they are in the left tail of the
residual distribution, indicating that the Kims ran faster than one
would have expected.  If one uses half marathon split times as a
measure of marathon talent, as in Figure
\ref{fig:studentizedresiduals_half}, a similar conclusion is reached.
Neither Hahner twin had a finishing time that was particularly unusual
given her half marathon split, and this applies to the Kim twins as
well.

@@

Figure \ref{fig:secondsbehind} provides a sense of the lockstep nature
of the Kim twins, and it describes each Olympic runner's status at
various split times on the marathon course.  Each dot in the figure
depicts a recorded split and the number of seconds each runner was
behind the race leader at the time.  There are more dots at earlier
splits due to subsequent DNFs.  The usual color scheme applies here,
and the Estonian triplet DNF prior to the half marathon split is
evident. As noted in the figure, Lisa Hahner was ahead of her sister
through 15 kilometers at which point Anna took the lead.  Figure
\ref{fig:secondsbehind} contains two red dots representing the North
Korean Kim twins, but this is not visually apparent because the Kim
twins had identical split times during the entire marathon.

@@

An important point regarding our simulations is the way that we
handled DNFs. As our description above indicates, we assumed that DNF
probabilities are the same for all runners and that the likelihood of
a DNF is a not function of a runner's anticipated marathon finishing
time. The rug marks in Figure \ref{fig:45degreeplot} suggest that
runners with better personal best times may be more likely to DNF than
other runners, all things equal. We suspect that this occurs because
some better runners may expend excessive energy trying to achieve a
good result in the marathon and in so doing injure or exhaust
themselves; lesser runners, in contrast, may be content to finish
respectably. Regardless of the validity of this conjecture, Figure
\ref{fig:45degreeplot} shows that the Hahner twins are representative
of the sort of runners who DNFed in Rio. Since our simulated Hahner
statistics are conditioned on both Hahner twins finishing, it follows
that they are conservative. The fact that both women finished the
marathon was notable in and of itself, and by discounting the
possibility of a Hahner DNF we are giving the benefit of the doubt to
the Hahner twins.
